### 2023-08-31 paper/markdown/supplementary_material.Rmd

---
title: "Three distinct patterns of mental health response following accidents in mountain sports – a follow-up study of individuals treated at a tertiary trauma center"
subtitle: "Supplementary Material"
author: "Psychiatry Study Team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx

bibliography: ptsd_biblio.bib
csl: frontiers_medical.csl

header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      dev = "png", 
                      dpi = 600, 
                      tab.cap.pre = 'Supplementary Table ', 
                      tab.cap.sep = " ", 
                      tab.cap.style = 'Table Heading')

set_flextable_defaults(font.family = 'Cambria', font.size = 9)

```

\newpage

# Supplementary Methods

## Software

Data analysis and transformation was accomplished with R version 4.2.3. 
The study dataset was imported from an SPSS file with the _foreign_ package [@RCoreTeam]. 
Tabular data were handled with the _tidyverse_ package bundle [@Wickham2019], and the packages _rlang_ [@Henry2022] and [_trafo_](https://github.com/PiotrTymoszuk/trafo). 
Text variables were handled with _stringi_ [@Gagolewski2021].

For distribution testing (normality, variance, Gini index), the packages [_ExDA_](https://github.com/PiotrTymoszuk/ExDA), _rstatix_ [@Kassambara2021] and _DescTools_ [@Signorell2022] were utilized. 
Consistency of psychometric tools was assessed with the packages _psych_ [@Revelle2022], [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) and [_clustTools_](https://github.com/PiotrTymoszuk/clustTools). 
Clustering tendency was investigated with _factoextra_ [@Kassambara2020] and [_clustTools_](https://github.com/PiotrTymoszuk/clustTools). 
Statistical hypothesis testing and correlation analyses were accomplished with [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) and _rstatix_ [@Kassambara2021].

For semi-supervised clustering, diagnostic and performance of the clustering analysis, the package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools) with implementation of algorithms and distance measures from _factoextra_, _cluster_ and _philentropy_ [@Schubert2019; @Drost2018; @Kassambara2020] was utilized. 
UMAP (uniform manifold approximation and projection) was done with the _umap_ and [_clustTools_](https://github.com/PiotrTymoszuk/clustTools) packages [@Konopka2022; @McInnes2018]. 
Cross-distances between the training and test subsets of the study cohort and cross-distances between clusters in the training and test subsets of the study cohort were computed with _proxy_ [@Meyer2022].

Multi-variable classifiers of cluster assignment were developed with the following algorithms: 
random forests [@Breiman2001; @Wright2017], 
neural network [@Ripley2014], 
support vector machines with radial kernel [@Weston1998; @Karatzoglou2004], 
recursive partitioning [@Therneau2022; @Breiman2017], 
conditional random forests [@Hothorn2006; @Strobl2007; @Hothorn2022], 
shrinkage discriminant analysis [@Ahdesmaki2010; @Ahdesmaki2022], 
and elastic net multinomial regression [@Zou2005; @Friedman2010]. 
For tuning, training, prediction and assessment of performance of the classifiers, the packages _caret_ [@Kuhn2008] and [_caretExtra_](https://github.com/PiotrTymoszuk/caretExtra) were used.

Results were visualized with _ggplot_ (bar plots, box plots, heat maps of cross-distances, scatter plots) [@Wickham2016], _plotroc_ [@Sachs2017], [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) (violin, stack and ribbon plots) and [_clustTools_](https://github.com/PiotrTymoszuk/clustTools) (cluster quality control plots, heat maps of clustering features, distance heat maps, scatter plots of UMAP layouts) and _ComplexUpset_ (visualization of overlap with upset plots) [@Krassowski2021]. 
Figures were created with the packages _cowplot_ [@Wilke2019] and _patchwork_ [@Pedersen2023]. 
Tables were generated with _flextable_ [@Gohel2022]. 
The manuscript and Supplementary Material were written in the _rmarkdown_ environment [@Allaire2022] with the package _bookdown_ [@Xie2016]. 
Figures, tables and R expressions in the markdown documents were managed with the development package [_figur_](https://github.com/PiotrTymoszuk/figur). 
The markdown documents were rendered with the _knitr_ [@Xie2022] and _bookdown_ [@Xie2016] packages and with [`author-info-blocks.lua`](https://github.com/pandoc/lua-filters/blob/master/author-info-blocks/author-info-blocks.lua) and [`scholarly-metadata.lua`](https://github.com/pandoc/lua-filters/blob/master/scholarly-metadata/scholarly-metadata.lua) scripts by Albert Krewinkel and Jörn Krenzer.

## Data import and transformation

The study data set was imported from an SPSS file with raw study data (function `read.spss()`, package _foreign_). 
The list of extracted variables with their description is available in __Supplementary Table S\@ref(tab:tab-analysis-vars)__. 
A total of `r nrow(ptsd$dataset)` participants with the complete set of psychometric battery variables (__Supplementary Table S\@ref(tab:tab-mental-battery)__) were included in the analysis (__Figure 1__).

The psychometric battery (__Supplementary Table S\@ref(tab:tab-mental-battery)__) consisted of German versions of assessment tools for 
anxiety (GAD-7: 7-item general anxiety disorder scale) [@Spitzer2006], 
depression (PHQ: patient health questionnaire, PHQ-9) [@Lowe2002; @Grafe2004], 
panic (PHQ-panic module) [@Lowe2002; @Grafe2004], 
persistent somatic symptoms (PHQ-15) [@Kroenke2002], 
resilience (RS13: 13-item resilience scale) [@Leppert2008], 
loss of sense of coherence (SOC-9L: Leipzig 9-item sense of coherence questionnaire) [@Schumacher2000], 
quality of life (EUROHIS-QOL 8: 8-item EUROHIS project quality of life scale) [@Schmidt2006], 
post-traumatic growth (PTGI: post-traumatic growth inventory) [@Tedeschi1996] and 
post-traumatic stress disorder (PCL-5 DSM-5: PTSD checklist for DSM-5) [@Bovin2016].

Clinically relevant symptoms of anxiety were defined as GAD-7 $geq$ 10, 
clinically relevant symptoms of depression were defined as PHQ-9 $\geq$ 11 [@Manea2012], 
significant persistent somatic symptoms were defined as PHQ-15 $\geq$ 11 [@Kroenke2002]. 
Resilience classes were defined as follows: low: RS-13 0 - 65, moderate: 66 - 72, high: $\geq$ 73 [@Leppert2008].

Items of the EUROHIS QOL 8 tool were scored as with 1 - 5 likert scales (1: extreme concerns, no satisfaction at all, 5: no concerns/full satisfaction) with each item representing a single domain of quality of life (quality of life, health, energy, finances, activity, self-esteem, relationship and housing). 
The total EUROHIS QOL score was defined as the arithmetic mean of all items [@Schmidt2006].

Separate scores were computed for each domain of the PTGI tool (I: relations, II: possibilities, III: personal strength, IV: spiritual strength, V: life appreciation) with each item scores as 0: none, 1: very little, 2: little, 3: moderate, 4: great, 5: extremely great. 
In addition, the total PTGI score was calculated as the sum of all items [@Tedeschi1996].

Separate scores were calculated for domains B, C, D and E of the PCL-5 DSM-5 tool along with the total score being the sum of all items. 
Each PCL-5 DSM-5 item was scored as 0: not at all, 1: a little bit, 2: moderate, 3: quite a bit and 4: extremely. 
Participants positive for the B domain or C domain PTSD symptoms were identified by at least one item per domain scored with 'moderate' or higher. 
Participants positive for the D or E domain PTSD symptoms were identified by at least two items per domain scored with 'moderate' or higher. 
Significant PTSD symptoms where assumed in participants screened positive for at least one of the B, C, D or E PCL-5 DSM-5 domains [@Bovin2016].

Traumatic events prior to the sport accident were assessed with the DIA-X tool (Diagnostic Expert System) [@Maercker2005]. 
Direct personal experience or being a witness of a traumatic event specified by the DIA-X questionnaire or by an additional yes/no item ('other traumatic events') was scored as 1. 
Prior traumatic event was assumed with at least one DIA-X item scored with 1. 
Prior sport accidents, 
flashbacks during mountain sport activity, 
confusion during mountain sport activity, 
self-reported need for psychological support, 
psychological/psychiatric support/therapy after the accident 
and presence of persistent physical health consequences after the accident 
were surveyed as single yes/no items. 
Flashbacks frequency during mountain sport activity was assessed in the following categories: none, more than one per year and more than one per month. 
Smoking was surveyed as a single yes/no question. 
Alcohol use was investigated with the CAGE tool with $\geq$ 2 points indicative of problematic alcohol consumption [@OBrien2008]. 
Data on the type of the accident date and daytime, accident mountain sport, injury diagnosis, injured body regions, injury severity, hospital treatment, surgery and number of ICD-10 surgical diagnoses  were extracted from electronic patient's records. 
Injury severity was assessed with the abbreviated injury scale (AIS) [@Gennarelli2006].

Additional information on study variables and their stratification schemes are presented in __Supplementary Table S\@ref(tab:tab-analysis-vars)__.

## Consistency of psychometric tools and power analysis

Consistency of psychometric tools was assessed by factor analysis and McDonald's $\omega$ (function `omega()`, package _psych_) [@McDonald1999; @Revelle2022]. 
The number of latent factors for calculation of $\omega$ was identified by inspection of loadings determined by factor analysis (function `reduce_data()`, [_clustTools_](https://github.com/PiotrTymoszuk/clustTools)) [@Bartlett1937]. 
All psychometric tools used in the study except for the stress PSS-4 scale [@Cohen1983] and the BRCS resilience tool [@Sinclair2004] exhibited good-to-excellent consistency with $\omega$ > 0.8 (__Supplementary Table S\@ref(tab:tab-consistency)__). 
The poorly performing PSS-4 and BRCS scales were excluded from the further analysis.

To find the optimal size of a training subset of the study cohort, random subsamples of the study dataset of varying sizes were investigated (`r paste(pow$sizes, collapse = ', ')` observations, 50 random draws per subsample size). 
For each random subsample, clustering tendency was assessed by Hopkins statistic (function `get_clust_tendency()`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools)). 
For n = 250 observations, the median Hopkins statistic was 
`r pow$hopkins %>% filter(size == 250) %>% .$hopkins_stat %>% median %>% signif(2)` (interquartile range: 
`r pow$hopkins %>% filter(size == 250) %>% .$hopkins_stat %>% quantile(probs = c(0.25, 0.75)) %>% signif(2) %>% paste(collapse = ' - ')`) indicative of good clustering tendency. 
The Hopkins statistic value for this subset was comparable with the value for the entire subset 
(`r signif(pca$clust_tendency$data$hopkins_stat, 2)`). 
For these reasons, 3/4 of the entire dataset or n = 
`r count(ptsd$dataset, partition)$n[1]` was considered as the adequate size of the training subset of the study cohort for a reproducible clustering analysis.

## Training/test subset definition

The study participants were assigned to the training and test subsets with the 3:1 size ratio (__Figure 1__). 
To this end, `r nrow(part$dists)` random splits of the datasets were generated. 
The subset assignment scheme was chosen with the possibly smallest differences in sociodemographic, medical history, clinical and accident- and injury-related variables assessed by Gower distance between the training and test subsets (function `dist()`, package _proxy_) [@Meyer2022].

## Statistical hypothesis testing, effect size and multiple testing correction

Differences in numeric variables were assessed with Mann-Whitney test with r effect size statistic or Kruskal-Wallis test with $\eta^2$ effect size statistic for two and more than two analysis groups, respectively. 
Differences in frequency of categories of qualitative variables between analysis groups were investigated with $\chi^2$ test with Cramer's V effect size statistic. 
P values were corrected for multiple testing with the false discovery rate method separately for each analysis task (e.g. comparison of clusters) [@Benjamini1995]. 
Effects with p < 0.05 following the false discovery rate adjustment were considered significant. 
Effect size of accuracy of predicted cluster assignment by machine learning classifiers was assessed by Cohen's $\kappa$ inter-rater reliability statistic [@Cohen1960; @McHugh2012]. 
Intervals of effect sizes were defined as follows [@McHugh2012; @Field2013; @Cohen2013]:

* r statistic: weak: < 0.3, moderate: 0.3 - 0.5, large: $\geq$ 0.5

* Cramer's V statistic: weak: < 0.3, moderate: 0.3 - 0.5, large: $\geq$ 0.5

* $\eta^2$ statistic: weak: < 0.06, moderate: 0.06 - 0.14, large: $\geq$ 0.14

* Cohen's $\kappa$ reliability: none: < 0.2, minimal: 0.2 - 0.4, weak: 0.4 - 0.6, moderate: 0.6 - 0.8, strong: $\geq$ 0.8

## Assessment of selection bias

To assess the possible selection bias, demographic, socioeconomic, clinical, accident- and recovery-related parameters were compared between individuals excluded due to missingness of psychometric data or denying survey participation and participants included in the analysis (__Figure 1__). 
Categorical variables were compared by $\chi^2$ test with Cramer V effect size statistic and numeric variables were compared by Mann-Whitney test with r effect size statistic 
(function `compare_variables()`, package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA)). 
Potential differences between the training and test subset of the study cohort (__Figure 1__) were analyzed in an analogical way. 
Significant (p < 0.05) differences between the included/excluded participants are presented in __Supplementary Table S\@ref(tab:tab-no-response)__ and __S\@ref(tab:tab-missing)__.

## Semi-supervised clustering

Observations of the training subset of the study cohort were subjected to clustering in respect to the numeric psychometric scores (__Supplementary Table S\@ref(tab:tab-mental-battery)__). 
The score values were normalized and median-centered (function `center_data()`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools)). 
For clustering, the PAM (partition around medoids) algorithm [@Schubert2019] with the cosine distance measure between observations [@Drost2018] was employed. 
The clustering object was constructed with the function `kcluster()` from the  [_clustTools_](https://github.com/PiotrTymoszuk/clustTools) package.
The choice of the clustering algorithm was motivated by its good explanatory performance measured by the fraction of explained clustering variance (ratio of between-cluster sum of squares to total sum of squares, method `var()`, [_clustTools_](https://github.com/PiotrTymoszuk/clustTools)) and superior reproducibility in 10-fold cross-validation [@Lange2004] (cluster assignment in the folds by an inverse distance weighted 27-nearest neighbors classifier, method `cv()`, [_clustTools_](https://github.com/PiotrTymoszuk/clustTools)) in a comparison with several other clustering algorithms presented in __Supplementary Figure S\@ref(fig:fig-clust-dev)A__. 
The number of cluster was chosen based on the bend of the curve of within-cluster sum of squares and the peak of mean silhouette statistic (method `plot()`, package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA)) [@Rousseeuw1987; @Kassambara2020]. 
By this means, three mental clusters were defined: the neutral, PTG (post-traumatic growth) and PTS (post-traumatic stress) cluster (__Supplementary Figure S\@ref(fig:fig-clust-dev)B__).

Assignment of the training subset observations to the mental clusters was accomplished with an inverse distance weighted 27-nearest neighbor classifier. 
This semi-supervised procedure yielded clustering structures with similar fractions of explained variance in the training (V = `r signif(semi_clust$variance$variance[1], 2)`) and test subset (V = `r signif(semi_clust$variance$variance[2], 2)`). 
Comparably good separation of the clusters could be discerned in the training and test subsets by 
a visual analysis of UMAP layouts and 
pairwise distance heat maps (__Supplementary Figure S\@ref(fig:fig-semi-clust)__). 
Distribution of the cluster frequency was similar in the training and test subset (__Supplementary Figure S\@ref(fig:fig-cosine)__). 
An analysis of cosine cross-distances between the clusters in the training and test subsets revealed far higher similarity of the corresponding mental clusters (i.e. neutral vs neutral, PTG vs PTG, PTS vs PTS) as compared with similarity of non-analogous clusters (e.g. neutral vs PTG) (distances computed with the function `dist()`, package _proxy_; __Supplementary Figure S\@ref(fig:fig-cosine)B__). 
Finally, quality of semi-supervised clustering was investigated by comparison of normalized levels of the clustering variables in the training and test subset by Kruskal-Wallis test with $\eta^2$ effect size statistic (function `compare_variables()`, package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA); __Figure 2__, __Supplementary Figure S\@ref(fig:fig-feat-hm)__, __Supplementary Table S\@ref(tab:tab-clust-fct)__).

Differences in frequencies of mental disorder symptoms as well as demographic, socioeconomic, clinical, accident- and recovery-related factors between the mental clusters were assessed in the entire cohort by $\chi^2$ test with Cramer V effect size statistic and by Kruskal-Wallis test with $\eta^2$ effect size statistic for categorical and numeric variables, respectively (function `compare_variables()`, package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA); __Supplementary Table S\@ref(tab:tab-clust-bcg)__).

## Cluster assignment classifiers

Two types of multi-parameter machine learning classifiers of the mental cluster assignment were developed in the training subset:

1. models employing candidate early predictors of cluster assignment, i.e. demographic, socioeconomic, medical history and accident-related explanatory factors available during acute medical management of the accident

2. models including additionally recovery-related predictors such as persistent physical health consequences, flashbacks or cautious behavior during sport

Of note, psychometric variables used for definition of the mental clusters as well as mental disorder symptoms, presence and frequency of flashbacks were excluded from the explanatory variable sets. 
The explanatory variables are listed in __Supplementary Table S\@ref(tab:tab-mod-variables)__.

The models employed the following algorithms: 
canonical random forests [@Breiman2001; @Wright2017], 
regularized neural networks with a single hidden layer [@Ripley2014], 
support vector machines with radial kernel [@Weston1998; @Karatzoglou2004], 
recursive partitioning [@Therneau2022; @Breiman2017], 
shrinkage discriminant analysis [@Ahdesmaki2010; @Ahdesmaki2022], 
conditional random forest [@Hothorn2006; @Strobl2007; @Hothorn2022], 
and elastic net multinomial regression [@Zou2005; @Friedman2010]. 
The optimal values of the algorithms' parameters were found by 10-fold cross-validation-based tuning with the maximal value of Cohen's $\kappa$ [@Cohen1960] as the tuning criterion. 
The tuning and fitting ('training') in the training subset of the study cohort was done with the wrapper function `train()` provided by the package _caret_ [@Kuhn2008]. 
For the random forest and conditional random forest algorithms, 1000 random trees each were constructed. 
Test statistic, p values, number of splits and other parameters of the conditional forest models were controlled with the `ForestControl` object returned by the convenience wrapper `cforest_unbiased()` provided by the _party_ package [@Hothorn2022]. 
The optimal algorithm parameter sets are listed in __Supplementary Table S\@ref(tab:tab-tuning)__. 
Predictions of the cluster assignment in the test subset of the study cohort were obtained with the `predict()` method from the [_caretExtra_](https://github.com/PiotrTymoszuk/caretExtra) package.

Overall accuracy and Cohen's $\kappa$ statistics [@Cohen1960] were computed with the `summary()` method from the [_caretExtra_](https://github.com/PiotrTymoszuk/caretExtra) package. 
Brier scores [@Brier1950] were computed with the following formula:

$$BS = \frac{1}{N}\times\Sigma_{i = 1}^{N}\Sigma_{c = 1}^R(f_{ic} - o_{ic})^2$$
where $f_{tc}$ is the output probability of assignment of the i-th observation to the c-th cluster, 
$o_{ic}$ is the numeric-coded actual assignment of the i-th observation to the c-th cluster, 
$R$ is the total cluster number and 
$N$ is the total observation number. 
Brier skill scores [@Brier1950] were computed with the following formula:

$$BSS = 1 - \frac{BS_{model}}{BS_{random}}$$

where $BS_{model}$ is the Brier score of the model and 
$BS_{random}$ is the Brier score obtained for a purely random cluster assignment. 
Brier scores, Brier skill scores, sensitivity and specificity of the PTS cluster assignment were computed with in-house developed R functions. 
Performance statistics for the training subset, 10-fold cross-validation and test subset of the study cohort are listed in __Supplementary Table S\@ref(tab:tab-early-classifiers)__ and __Supplementary Table S\@ref(tab:tab-full-classifiers)__ for the classifiers employing the early and full predictor set, respectively.

Variable importance statistic specific for the machine learning algorithm were extracted from the _caret_ models with the `varImp()` function from package _caret_ [@Kuhn2008]. 
They were: permutation importance for the random forest and conditional forest algorithm [@Hothorn2022; @Breiman2001], 
connection weight importance statistic for the neural network [@Garson1991], 
sum reduction in classification error attributed to each variable at each split for recursive partitioning [@Breiman2017], 
and linear model coefficient $\beta$ for elastic net regression [@Zou2005]. 
For support vector machines and discriminant analysis, the variable importance was computed based on area under the ROC curve for single predictors for discrimination between the cluster pairs [@Kuhn2008].

## Data and code availability

An RData file with anonymized patient data will be made available upon request to the corresponding author. 
The study analysis pipeline is available at https://github.com/PiotrTymoszuk/mental_accident.

\newpage

# Supplementary Tables

```{r tab-analysis-vars, tab.cap = 'Variables used in the analysis pipeline.'}

suppl_tables$analysis_vars[1, 1] %>% 
  mutate(`Variable name in R` = '') %>% 
  set_names(' ') %>% 
  flextable %>% 
  set_widths(16)

```

\newpage

```{r tab-mental-battery, tab.cap = 'Mental health assessment battery.'}

flextable::flextable(suppl_tables$mental_battery) %>% 
  set_widths(c(2.3, 7, 10)) %>% 
  merge_v(1) %>% 
  footnote(1, 2:3, 
           value = as_paragraph('GAD-7: 7-item general anxiety disorder scale; PHQ: patient health questionnaire; EUROHIS-QOL 8: 8-item EUROHIS project quality of life scale; SOC-9L: Leipzig 9-item sense of coherence questionnaire; RS13: 13-item resilience scale; PCL-5 DSM-5: PTSD checklist for DSM-5; PTGI: post-traumatic growth inventory'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  theme_vanilla

```

\newpage

```{r tab-consistency, tab.cap = "Consistency of the psychometric tools used in the study measured by McDonald's omega."}

flextable::flextable(suppl_tables$consistency) %>% 
  set_widths(c(3.5, 3, 3)) %>% 
   footnote(1, 1, 
           value = as_paragraph('GAD-7: 7-item general anxiety disorder scale; PHQ: patient health questionnaire; EUROHIS-QOL 8: 8-item EUROHIS project quality of life scale; SOC-9L: Leipzig 9-item sense of coherence questionnaire; RS13: 13-item resilience scale; PCL-5 DSM-5: PTSD checklist for DSM-5; PTGI: post-traumatic growth inventory'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  theme_vanilla

```

\newpage

```{r tab-no-response, tab.cap = 'Significant differences between patients who did not respond to the study invitation and the analyzed study participants. Numeric variables are presented as medians with interquartile ranges (IQR). Categorical variables are presented as percentages and counts within the complete observation set.'}

flextable::flextable(suppl_tables$no_response) %>% 
  set_widths(c(3.2, 5.8, 5.8, 2.5, 2.2)) %>% 
   footnote(1, 1, 
           value = as_paragraph('AIS: abbreviated injury scale'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 4:5, 
           value = as_paragraph('Numeric variables: Mann-Whitney test with r effect size statistic; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic. P values corrected for multiple testing with the false discovery rate method.'), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla

```

\newpage

```{r tab-missing, tab.cap = 'Significant differences between the study survey responders excluded from analysis due to missingness of psychometric data and the analyzed study participants. Numeric variables are presented as medians with interquartile ranges (IQR). Categorical variables are presented as percentages and counts within the complete observation set.'}

flextable::flextable(suppl_tables$missing) %>% 
  set_widths(c(3.2, 5.8, 5.8, 2.5, 2.2)) %>% 
  footnote(1, 1, 
           value = as_paragraph('AIS: abbreviated injury scale; K: 1000 Euro; EUR: Euro.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 4:5, 
           value = as_paragraph('Numeric variables: Mann-Whitney test with r effect size statistic; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic. P values corrected for multiple testing with the false discovery rate method.'), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla

```

\newpage

```{r tab-train-test, tab.cap = 'Significant differences between the training and test subset of the study cohort. Numeric variables are presented as medians with interquartile ranges (IQR). Categorical variables are presented as percentages and counts within the complete observation set.'}

flextable::flextable(suppl_tables$train_test) %>% 
  set_widths(c(4, 4, 4, 2.5, 2.2)) %>% 
  footnote(1, 1, 
           value = as_paragraph('RS13: 13-item resilience scale; EUROHIS-QOL 8: 8-item EUROHIS project quality of life scale.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 4:5, 
           value = as_paragraph('Numeric variables: Mann-Whitney test with r effect size statistic; categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic. P values corrected for multiple testing with the false discovery rate method.'), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla

```

\newpage

```{r tab-clust-fct, tab.cap = 'Differences in psychometric clustering factors between the mental clusters. Numeric variables are presented as medians with interquartile ranges (IQR). Statistical significance was determined by false discovery rate-corrected Kruskal-Wallis test with eta-square effect size statistic. The table is available in a supplementary Excel file.'}

suppl_tables$clust_fct[1, 1] %>% 
  mutate(Variable = '') %>% 
  set_names(' ') %>% 
  flextable %>% 
  set_widths(16)

```

\newpage

```{r tab-clust-bcg, tab.cap = 'Significant differences in demographic, socioeconomic, clinical, accident- and recovery-related factors, and mental disorder symptoms between the mental clusters in the entire cohort. Numeric variables are presented as medians with interquartile ranges (IQR). Categorical variables are presented as percentages and counts within the complete observation set.'}

flextable::flextable(suppl_tables$clust_bcg) %>% 
  set_widths(c(3, rep(4.3, 3), 2.5, 2.3)) %>% 
  footnote(1, 1, 
           value = as_paragraph('AIS: abbreviated injury score; PTSD: post-traumatic stress disorder; RS13: 13-item resilience scale; GAD-7: 7-item general anxiety disorder scale; PHQ: patient health questionnaire.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 5:6, 
           value = as_paragraph('Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic. Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic. P values were corrected for multiple testing with the false discovery rate method.'), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla

```

\newpage

```{r tab-mod-variables, tab.cap = 'Sets of explanatory factors used for modeling of the mental cluster assignment.'}

flextable::flextable(suppl_tables$mod_variables) %>% 
  set_widths(c(2.5, 15.5)) %>% 
  footnote(1, 1, 
           value = as_paragraph('early predictors: variables available during acute medical management of the accident; full predictor set: variables available during acute medical management of the accident and during follow-up.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  theme_vanilla

```

\newpage

```{r tab-tuning, tab.cap = 'The optimal combinations of machine learning algorithm parameters found in 10-fold cross-validation of the training subset of the study cohort.'}

flextable::flextable(suppl_tables$tuning) %>% 
  set_widths(c(2, 2.5, 8.5)) %>% 
  merge_v(1) %>%
  footnote(1, 1, 
           value = as_paragraph('early predictors: variables available during acute medical management of the accident; full predictor set: variables available during acute medical management of the accident and during follow-up.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 2, 
           value = as_paragraph('RF: random forest; NNet: neural network with a single hidden layer; SVM/radial: support vector machines with radial kernel; RPart: recursive partitioning; SDA: shrinkage discriminant analysis; cForest: conditional random forest; Elastic Net: elastic net multinomial regression.'), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla

```

\newpage

```{r tab-early-classifiers, tab.cap = 'Performance statistics of machine learning classifiers at predicting the mental cluster assignment. Models employing early predictors available during acute medical management of the accident.'}

flextable::flextable(suppl_tables$early_classifiers) %>% 
  set_widths(c(2.3, 2.5, rep(2.3, 6))) %>% 
  merge_v(1) %>% 
  footnote(1, 1, 
           value = as_paragraph('RF: random forest; NNet: neural network with a single hidden layer; SVM/radial: support vector machines with radial kernel; RPart: recursive partitioning; SDA: shrinkage discriminant analysis; cForest: conditional random forest; Elastic Net: elastic net multinomial regression.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 2, 
           value = as_paragraph('CV: cross-validation'), 
           part = 'header',  
           ref_symbols = 'b') %>% 
  footnote(1, 6, 
           value = as_paragraph('Brier Skill Score comparing the given classifier with the purely random cluster assignment.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  theme_vanilla

```

\newpage

```{r tab-full-classifiers, tab.cap = 'Performance statistics of machine learning classifiers at predicting the mental cluster assignment. Models employing the full predictor set available during acute medical management of the accident and follow-up.'}

flextable::flextable(suppl_tables$full_classifiers) %>% 
  set_widths(c(2.3, 2.5, rep(2.3, 6))) %>% 
  merge_v(1) %>% 
  footnote(1, 1, 
           value = as_paragraph('RF: random forest; NNet: neural network with a single hidden layer; SVM/radial: support vector machines with radial kernel; RPart: recursive partitioning; SDA: shrinkage discriminant analysis; cForest: conditional random forest; Elastic Net: elastic net multinomial regression.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 2, 
           value = as_paragraph('CV: cross-validation'), 
           part = 'header',  
           ref_symbols = 'b') %>% 
  footnote(1, 6, 
           value = as_paragraph('Brier Skill Score comparing the given classifier with the purely random cluster assignment.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  theme_vanilla

```

\newpage

# Supplementary Figures

```{r fig-parts-ptsd, fig.width = figur::convert(suppl_figures$parts_ptsd, to = 'in')$w, fig.height = figur::convert(suppl_figures$parts_ptsd, to = 'in')$h, fig.cap = 'Injured body regions and symptoms of PTSD in the study cohort.'}

suppl_figures$parts_ptsd$plot

```

__Supplementary Figure S\@ref(fig:fig-parts-ptsd). Injured body regions and symptoms of PTSD in the study cohort.__

_(A) Distribution of injured body regions presented in a bar plot. The number of complete observations is indicated in the plot caption._

_(B) Scores of particular domains of the PCL-5 tool evaluating symptoms of post-traumatic stress disorder (PTSD) presented in violin plots. Single observations are visualized as points. Medians and interquartile ranges are represented by red diamonds and whiskers. The number of complete observations is indicated in the plot caption._

_(C) Frequency and overlap of clinically relevant PTSD symptoms captured by particular domains of the PCL-5 scale presented in an upset plot. Percentages of participants screened positive for the domain B, C, D or E PTSD symptoms are displayed in as red bars. Percentages of participants with single and overlapping domain B, C, D or E PTSD symptoms are presented as green bars. The number of complete observations is indicated in the plot caption._

\newpage

```{r fig-qol-ptg, fig.width = figur::convert(suppl_figures$qol_ptg, to = 'in')$w, fig.height = figur::convert(suppl_figures$qol_ptg, to = 'in')$h, fig.cap = 'Scores of quality of life and post-traumatic growth in the study cohort.'}

suppl_figures$qol_ptg$plot

```

__Supplementary Figure S\@ref(fig:fig-qol-ptg). Scores of quality of life and post-traumatic growth in the study cohort.__

_(A) Scores of the domains of the EUROHIS project 8-item quality of life scale (EUROHIS-QOL 8) presented in a bar plot. The number of complete observations is displayed in the plot caption._

_(B) Scores of the domains of the post-traumatic growth inventory (PTGI) scale presented in a violin plots. Single observations are visualized as points. Red diamonds with whiskers represent medians with interquartile ranges. The number of complete observations is indicated in the plot caption._

\newpage

```{r fig-clust-dev, fig.width = figur::convert(suppl_figures$clust_dev, to = 'in')$w, fig.height = figur::convert(suppl_figures$clust_dev, to = 'in')$h, fig.cap = 'Definition of the mental clusters in the training subset of the study cohort.'}

suppl_figures$clust_dev$plot

```

__Supplementary Figure S\@ref(fig:fig-clust-dev). Definition of the mental clusters in the training subset of the study cohort.__

_The mental clusters were defined in respect to psychometric scoring in the training subset of the study cohort by PAM (partition around medoids) with cosine distance between the observations._

_(A) Comparison of explanatory performance and reproducibility of various clustering algorithms in the training subset. The explanatory performance was measured as a fraction of explained clustering variance (ratio of between-cluster sum of squares to total sum of squares). The reproducibility was assessed by the rate of correct cluster assignment in 10-fold cross-validation (CV) with cluster assignment in the folds by an inverse distance weighted 27-nearest neighbors classifier. Note the superior stability of the PAM/cosine distance algorithm._

_(B) Determination of the cluster number by the bend of the within-cluster sum of squares curve and the peak mean silhouette statistic._

\newpage

```{r fig-semi-clust, fig.width = figur::convert(suppl_figures$semi_clust, to = 'in')$w, fig.height = figur::convert(suppl_figures$semi_clust, to = 'in')$h, fig.cap = 'Semi-supervised clustering.'}

suppl_figures$semi_clust$plot

```

__Supplementary Figure S\@ref(fig:fig-semi-clust). Semi-supervised clustering.__

_The mental clusters were defined in respect to psychometric scoring in the training subset of the study cohort by PAM (partition around medoids) with cosine distance between the observations. Assignment of the test subset observations to the mental clusters was done with the inverse distance weighted 27-nearest neighbors classifier. By this procedure, three mental clusters were identified: neutral, PTG (post-traumatic growth) and PTS (post-traumatic stress). Numbers of observations in the mental clusters are shown in the plot legend in (A)._

_(A) Observations in the training and test subsets of the study cohorts were subjected to two-dimensional UMAP (uniform manifold approximation and projection) in respect to the psychometric scores. UMAP layouts are shown in scatter plots. Points represent single samples. Point color codes for the cluster assignment._

_(B) Pairwise cosine distances between observations in the mental clusters of the training and test subsets presented in heat maps._

\newpage

```{r fig-cosine, fig.width = figur::convert(suppl_figures$cosine, to = 'in')$w, fig.height = figur::convert(suppl_figures$cosine, to = 'in')$h, fig.cap = 'Distribution of the mental clusters and cosine distances between the mental clusters in the training and test subset of the study cohort.'}

suppl_figures$cosine$plot

```

__Supplementary Figure S\@ref(fig:fig-cosine). Distribution of the mental clusters and cosine distances between the mental clusters in the training and test subset of the study cohort.__

_Study cohorts observations were assigned to the neutral, PTG (post-traumatic growth) and PTS (post-traumatic stress) clusters by semi-supervised clustering as presented in Supplementary Figure S\@ref(fig:fig-semi-clust)._

_(A) Percentages of observations in the training and test subsets of the study cohort in the mental clusters. Numbers of complete observations are indicated in the Y axis._

_(B) Cosine cross-distances between the mental clusters in the training and test subset of the study dataset visualized as a heat map. Cross-distances are indicated in the tiles._

\newpage

```{r fig-feat-hm, fig.width = figur::convert(suppl_figures$feat_hm, to = 'in')$w, fig.height = figur::convert(suppl_figures$feat_hm, to = 'in')$h, fig.cap = 'Levels of psychometric scores used for the cluster definition in the mental clusters.'}

suppl_figures$feat_hm$plot

```

__Supplementary Figure S\@ref(fig:fig-feat-hm). Levels of psychometric scores used for the cluster definition in the mental clusters.__

_Levels of psychometric scores used in clustering in the mental clusters. Normalized, median-centered score values (Z scores) are presented as heat maps. Numbers of observations in the clusters are displayed in the plot captions. PSS4: 4-item perceived stress scale; GAD-7: 7-item general anxiety disorder scale; PHQ: patient health questionnaire; EUROHIS-QOL 8: 8-item EUROHIS project quality of life scale; SOC-9L: Leipzig 9-item sense of coherence questionnaire; RS13: 13-item resilience scale; PCL-5 DSM-5: PTSD checklist for DSM-5; PTGI: post-traumatic growth inventory; PTSD: post-traumatic stress disorder._

\newpage

```{r fig-demo, fig.width = figur::convert(suppl_figures$demo, to = 'in')$w, fig.height = figur::convert(suppl_figures$demo, to = 'in')$h, fig.cap = 'Employment status, prior sport accidents, and accident details in the mental clusters.'}

suppl_figures$demo$plot

```

__Supplementary Figure S\@ref(fig:fig-demo). Employment status, prior sport accidents, and accident details in the mental clusters.__

_Distribution of employment status, frequency of sport accidents in the past, distribution of the accidents in the seasons, accident sport types, frequency of being alone during the accident and number of injured persons in the mental clusters. Statistical significance was determined by $\chi^2$ test with Cramer V effect size statistic. P values were corrected for multiple testing with the false discovery rate method. Percentages of variable's categories in the entire study cohort are presented in stack plots. Effect sizes and p-values are displayed in the plot captions. Numbers of observations in the clusters are presented in the X axes._

\newpage

```{r fig-importance-early, fig.width = figur::convert(suppl_figures$importance_early, to = 'in')$w, fig.height = figur::convert(suppl_figures$importance_early, to = 'in')$h, fig.cap = 'Variable importance metrics for the random forest, support vector machine, discriminant analysis, and conditional random forest algorithms. Predictors available during acute medical management of the accident.'}

suppl_figures$importance_early$plot

```

__Supplementary Figure S\@ref(fig:fig-importance-early). Variable importance metrics for the random forest, support vector machine, discriminant analysis, and conditional random forest algorithms. Predictors available during acute medical management of the accident.__

_The cluster assignment was modeled with demographic, medical history and accident-related explanatory factors available during acute medical management of the accident. Psychometric variables used for cluster definition, mental disorder symptoms, resilience classes as well as presence and frequency of flashbacks were excluded from the explanatory factor set. Variable importance metrics of machine learning algorithms with the best performance at predicting the mental cluster assignment in the test subset of the study cohorts were computed. Importance metrics for the top 20 most important variables are presented as bar plots._
_(A) Random forest, (B) support vector machines with radial kernel, (C) shrinkage discriminant analysis, (D) conditional random forest._

\newpage

```{r fig-full-class, fig.width = figur::convert(suppl_figures$full_class, to = 'in')$w, fig.height = figur::convert(suppl_figures$full_class, to = 'in')$h, fig.cap = 'Assignment of accident victims to the mental clusters based on explanatory factors available during acute medical management of the accident and long-term follow-up.'}

suppl_figures$full_class$plot

```

__Supplementary Figure S\@ref(fig:fig-full-class). Assignment of accident victims to the mental clusters based on explanatory factors available during acute medical management of the accident and long-term follow-up.__

_The cluster assignment was modeled with demographic, medical history, accident- and recovery-related explanatory factors available during acute medical management of the accident and follow-up. Psychometric variables used for cluster definition, mental disorder symptoms, resilience classes as well as presence and frequency of flashbacks were excluded from the explanatory factor set._

_(A) Accuracy of the predicted Cluster assignment and predictive performance of the modeling algorithms was assessed by overall cluster assignment accuracy, Cohen's $\kappa$ inter-rater accuracy metric (high values indicate good accuracy) and Brier score (low values indicate good performance) in the training and test subsets of the study cohort. Performance metrics are presented in scatter plots. Point size codes for the overall cluster assignment accuracy. Point color codes for the modeling algorithm. Numbers of complete observations are displayed in the plot captions._

_(B) Sensitivity (Se) and specificity (Sp) of detection of participants assigned to the PTS cluster (post-traumatic stress) investigated by receiver-operating characteristic in the training and test subset of the study cohort. Sensitivity and specificity values are indicated in the plots. Line color codes for the modeling algorithm. Numbers of complete observations and observations in the PTS cluster are indicated in the plot captions._

_RF: random forest; NNet: neural network with a single hidden layer; SVM/radial: support vector machines with radial kernel; RPart: recursive partitioning; SDA: shrinkage discriminant analysis; cForest: conditional random forest; Elastic Net: elastic net multinomial regression._

\newpage

```{r fig-importance-full, fig.width = figur::convert(suppl_figures$importance_full, to = 'in')$w, fig.height = figur::convert(suppl_figures$importance_full, to = 'in')$h, fig.cap = 'Variable importance metrics for the random forest, support vector machine, discriminant analysis, and conditional random forest algorithms. Predictors available during acute medical management of the accident and during the follow-up.'}

suppl_figures$importance_full$plot

```

__Supplementary Figure S\@ref(fig:fig-importance-full). Variable importance metrics for the random forest, support vector machine, discriminant analysis, and conditional random forest algorithms. Predictors available during acute medical management of the accident and during the follow-up.__

_The cluster assignment was modeled with demographic, medical history, accident- and recovery-related explanatory factors available during acute medical management of the accident and follow-up. Psychometric variables used for cluster definition, mental disorder symptoms, resilience classes as well as presence and frequency of flashbacks were excluded from the explanatory factor set. Variable importance metrics of machine learning algorithms with the best performance at predicting the mental cluster assignment in the test subset of the study cohorts were computed. Importance metrics for the top 20 most important variables are presented as bar plots._
_(A) Random forest, (B) support vector machines with radial kernel, (C) shrinkage discriminant analysis, (D) conditional random forest._

\newpage

# References